The problem and promise of scale in animal social multilayer networks

### 2020-02-10 paper/manuscript.Rmd
refs blah blah

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Social phenotypes are
increasingly linked to important fitness for individuals of various species,
including primates (e.g. Brent, Ruiz-Lambides, & Platt, 2017; Thompson, 2019),
dolphins (e.g. Stanton & Mann, 2012; Stanton, Gibson, & Mann, 2011), and birds
(e.g. McDonald, 2007; Royle, Pike, Heeb, Richner, & Kolliker, 2012). Both
temporal, seasons and stages of life, and social scales, types of interactions,
have emerged as important for understanding fitness consequences in the context
of social network positions (Almeling, Hammerschmidt, Sennhenn-Reulen, Freund, &
Fischer, 2016; Berger, Lemaître, Allainé, Gaillard, & Cohas, 2015; Brent,
Ruiz-Lambides, & Platt, 2017; Holekamp, Smith, Strelioff, Van Horn, & Watts,
2012). Multilayer networks provide a way to bridge these different scales so
that we can examine social phenotypes in more holistic ways rather than looking
at each scale separately as has traditionally been done.

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Cater_2015; @Farine_2015a].
Furthermore, it is a common trope that individuals must be associating to
interact, and many studies therefore assume that proximity is proxy for
interacting (Farine, 2015). For instance, baboons need to be in proximity in
order to groom each other. However, animals have certain interactions that do
not require proximity such as long-distance vocalization and olfactory signals
[@Carter_2015] With the advancement of biologging technology, we have
the ability to record social interactions that do not require proximity,
particularly long-distance communication networks with devices like microphone
arrays (Snijders & Naguib, 2017). These varying social scales can have different
fitness repercussions for individuals. Rhesus macaques (*Macaca mulatta*) had
higher survivorship when they had fewer but stronger, more stable social
associates and grooming partners rather than simply more associates (Ellis,
Snyder-Mackler, Ruiz-Lambides, Platt, & Brent, 2019). This finding highlights
that different qualities of social scales have lasting consequences.

Temporal scale is also important for understanding the repercussions of social
phenotypes. For instance, social phenotypes can change over ontogeny in ways
that can affect fitness (Berger et al., 2015; Brent et al., 2017). Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates (Tarka, Guenther, Niemelä, Nakagawa, & Noble, 2018). This variation can
make defining appropriate temporal scales for testing the fitness implications
of an individual’s social phenotype over different stages of ontogeny even more
difficult when they vary by individual. However, considering fine temporal
scales provide insight across developmental stages that may otherwise be missed
(Turner, Bills, & Holekamp, 2018).

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful simplification of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014]. Furthermore, multilayer networks provide a
framework to test hypotheses linking social phenotypes and fitness across
temporal scales. Because of the non-independent nature of social interactions
and associations, especially over time, it is hard to test hypotheses about
social dynamics with a suitable null control (Farine, 2017; Proskurnikov &
Tempo, 2017). Using the framework of multilayer networks, making layers of
different temporal periods provides a way to make appropriate null models to
test questions about the dynamics of an individual’s social network position
over multiple aggregated time periods.

On the other hand, picking the proper temporal scale is important but a
challenge because all layers in the multilayer network should reflect the same
time period, and the period should be long enough for each layer to have enough
data for robust networks (Farine, 2017). These considerations may be limiting
factors when answering questions that require timeframes that are relevant to
different individuals, for instance. In both mono- and multilayer networks, the
effect of scale on the relationship between an individual’s social network
position and fitness need to be considered based on the hypotheses being tested
and logistical constraints.

>>>>

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Social phenotypes are
increasingly linked to important fitness for individuals of various species,
including primates [@Brent_2017; @Thompson_2019],
dolphins [@Stanton_2012; @Stanton_2011], and birds
[@McDonald_2007; @Royle_2012]. Both
temporal, seasons and stages of life, and social scales, types of interactions,
have emerged as important for understanding fitness consequences in the context
of social network positions [@Almeling_2016, @Berger_2015, @Brent_2017, @Holekamp_2012]. Multilayer networks provide a way to bridge these different scales so
that we can examine social phenotypes in more holistic ways rather than looking
at each scale separately as has traditionally been done.

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Cater_2015; @Farine_2015a].
Furthermore, it is a common trope that individuals must be associating to
interact, and many studies therefore assume that proximity is proxy for
interacting [@Farine_2015a]. For instance, baboons need to be in proximity in
order to groom each other. However, animals have certain interactions that do
not require proximity such as long-distance vocalization and olfactory signals
[@Carter_2015] With the advancement of biologging technology, we have
the ability to record social interactions that do not require proximity,
particularly long-distance communication networks with devices like microphone
arrays [Snijders_2017]. These varying social scales can have different
fitness repercussions for individuals. Rhesus macaques (*Macaca mulatta*) had
higher survivorship when they had fewer but stronger, more stable social
associates and grooming partners rather than simply more associates [Ellis_2019]. This finding highlights
that different qualities of social scales have lasting consequences.

Temporal scale is also important for understanding the repercussions of social
phenotypes. For instance, social phenotypes can change over ontogeny in ways
that can affect fitness [@Berger_2015, @Brent_2017]. Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates [@Tarka_2018]. This variation can
make defining appropriate temporal scales for testing the fitness implications
of an individual’s social phenotype over different stages of ontogeny even more
difficult when they vary by individual. However, considering fine temporal
scales provide insight across developmental stages that may otherwise be missed
[@Turner_2018].

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful simplification of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014]. Furthermore, multilayer networks provide a
framework to test hypotheses linking social phenotypes and fitness across
temporal scales. Because of the non-independent nature of social interactions
and associations, especially over time, it is hard to test hypotheses about
social dynamics with a suitable null control [@Farine_2017, @Proskurnikov_2017]. Using the framework of multilayer networks, making layers of
different temporal periods provides a way to make appropriate null models to
test questions about the dynamics of an individual’s social network position
over multiple aggregated time periods.

On the other hand, picking the proper temporal scale is important but a
challenge because all layers in the multilayer network should reflect the same
time period, and the period should be long enough for each layer to have enough
data for robust networks [@Farine_2017]. These considerations may be limiting
factors when answering questions that require timeframes that are relevant to
different individuals, for instance. In both mono- and multilayer networks, the
effect of scale on the relationship between an individual’s social network
position and fitness need to be considered based on the hypotheses being tested
and logistical constraints.
